---
title: "Reproducible Research Course Project 1"
author: "Chester Sellers"
date: "5/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Reproducible Research Course Project 1

## load my librarys

```{r }
library(ggplot2)
library(dplyr)

```

## Loading and preprocessing the data ### Q1

```{r }
activity <- read.csv("C:/Users/casel/Desktop/Coursera/Reproducible_Research/repdata_data_activity/activity.csv")
```

## View our Data
```{r }
View(activity)

```

## View a portion of data and its structure###
 ```{r }
  head(activity)
```
 
## Check basic statistics of the data ###
 ```{r }
  summary(activity)
 
```  
## Check the column names
 ```{r }
  names(activity)
```
## basic pairs plot not really necesary
```{r }
  pairs(activity) 
```    
## What is mean total number of steps taken per day?###  
```{r } 
stepsDay <- aggregate(steps ~ date, activity, sum, na.rm=TRUE)
``` 
 
## View a portion of data and its structure###
```{r }
  head(stepsDay)
```

## Histogram of the total number of steps taken each day ## Q2
```{r }  
  g <- ggplot(stepsDay, aes(steps))
  g+geom_histogram(boundary=0, binwidth=2000, col="black",fill="blue")+ggtitle("Histogram Steps per Day")+xlab("Steps")+ylab("Frequency")
```


## Mean and Median of the total number of steps taken per day ## Q3

## Get the Mean of steps per day
```{r }
  MeanStepsPerDay <- mean(stepsDay$steps)
```
## Result of Mean
```{r }
  MeanStepsPerDay
```
## Get the Median of steps per day
```{r }
  MedianStepsPerDay <- median(stepsDay$steps)
```  
## Result of Median
```{r }
  MedianStepsPerDay
```
# What is the average daily activity pattern? ## Q4

## Variable with my Data SPT = Steps Pattern in Time
```{r } 
  SPT <- aggregate(steps~interval,data=activity,FUN=mean,na.action=na.omit)
```
## Plot the result of Average Steps per Time Interval
```{r } 
  g1 <- ggplot(SPT, aes(interval, steps))
  g1+geom_line(col="blue")+ggtitle("Average Steps per Time Interval")+xlab("Intervals")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## Variable with my Data MISteps = Maximum Interval Steps
```{r } 
  MISteps <- SPT[which.max(SPT$steps),]$interval
  MISteps  
```

#   Imputing missing values   ## Q5

## Calculate and report the total number of missing values in the dataset 
```{r }
  TMValues <- sum(is.na(activity$steps))
  TMValues
```
  
## Devise a strategy for filling in all of the missing values in the dataset.   
```{r }  
  activity$fillSteps <- ifelse(is.na(activity$steps), round(SPT$steps[match(activity$interval, SPT$interval)],0), activity$steps)
```
### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r }
  activityNoMissingData <- data.frame(steps=activity$fillSteps, interval=activity$interval, date=activity$date)
```
### Cheching top and bottom row of our data
```{r }
  head(activityNoMissingData)  
  tail(activityNoMissingData)
``` 
### Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
```{r }
  SPDNM <- aggregate(activityNoMissingData$steps, list(activityNoMissingData$date), FUN=sum)
``` 
### Checking my data and column names
```{r }  
  head(SPDNM)
``` 
### Assing new header column names
```{r }
  colnames(SPDNM) <- c("Date", "Steps")
```
### Checking my data and column names  
```{r }
  head(SPDNM)
``` 
### Make a histogram
```{r } 
  g3 <- ggplot(SPDNM, aes(Steps))
  g3+geom_histogram(boundary=0, binwidth=2500, col="black", fill="blue")+ggtitle("Histogram Steps per Day")+xlab("Steps")+ylab("Frequency")
``` 
  
## Calculate and report the mean and median total number of steps taken per day.
 
### Result of Mean
```{r }
  mean(SPDNM$Steps)
```
### Result of Median
```{r }  
  median(SPDNM$Steps)
```

###  Are there differences in activity patterns between weekdays and weekends? Q6
  
### Create variable with date in correct format
```{r } 
    activityNoMissingData$RealDate <- as.Date(activityNoMissingData$date, format = "%Y-%m-%d")
    head(activityNoMissingData)
 ```   
### Create a variable with weekdays name
```{r } 
    activityNoMissingData$weekday <- weekdays(activityNoMissingData$RealDate)
    head(activityNoMissingData)
```    
### create a new variable indicating weekday or weekend
```{r } 
      activityNoMissingData$DayType <- ifelse(activityNoMissingData$weekday=='Saturday' | activityNoMissingData$weekday=='Sunday', 'weekend','weekday')
```    
### View and check our data
```{r } 
    head(activityNoMissingData)
    tail(activityNoMissingData)
```
    
### Make a panel plot containing a time series plot, type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)  
  
### Create variable with steps per time across week days // weekend days
```{r } 
    SPTDT <- aggregate(steps~interval+DayType,data=activityNoMissingData,FUN=mean)
```
### Time variable
```{r } 
    SPTDT$time <- SPTDT$interval
```
### draw the line plots
```{r } 
    g4 <- ggplot(SPTDT, aes(time, steps))
    g4+geom_line(col="blue")+ggtitle("Average Steps per Time Interval (Weekdays / Weekends Comparison)")+xlab("Time")+ylab("Steps")+facet_grid(DayType ~ .)
```







